---
title: "STAT_415 - Regression_Analysis"
author: "Naomi Carrigg"
date: "10/31/24"
output:
  pdf_document: 
    toc: no
    toc_depth: 2
    number_sections: no
---

```{r, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(lmtest)
library(car)
library(glmnet)

reduced_data <- readRDS('results.rds')
AirQuality_data_aqi_reduced <- readRDS('airquality_data_aqi_red.rds')
```

# Model Fitting
## Model 1: Predicting Carbon Monoxide with Nitrogen Dioxide and season using AQI
```{r}
model_1_data <- na.omit(subset(AirQuality_data_aqi_reduced, 
                               select = c("Season", "Month", "Year", 
                                          "SITE_NAME", "nitrogen_dioxide_no", 
                                          "carbon_monoxide")))
model_1 <- lm(carbon_monoxide ~ Season + nitrogen_dioxide_no, 
              data = model_1_data)
summary(model_1)
```

## Hypothesis test for Overall Significance

```{r}
null <- lm(carbon_monoxide ~ 1 , data = model_1_data)
anova(null,model_1)
```
Model is significant overall.

## Model 2: Predicting Nitrogen Dioxide AQI using all other predictors. Attempt to see if relationship exists in the opposite direction and whether location has an effect.
```{r}
model_2_data <- na.omit(subset(AirQuality_data_aqi_reduced, 
                               select = c("Season", "Month", "Year", 
                                          "SITE_NAME", "nitrogen_dioxide_no")))

model_2_red <- lm(nitrogen_dioxide_no ~ SITE_NAME + Season + Month + Year, 
                  data = model_1_data)

model_2_full <- lm(nitrogen_dioxide_no ~ SITE_NAME + Season + Month + Year + 
                     carbon_monoxide, data = model_1_data)

anova(model_2_red, model_2_full)
```

```{r}
summary(model_2_full)
summary(model_2_red)
```

Carbon monoxide can be dropped from the model when we include season, month, year, and Site_Name. Now we perform best subset selection on this model

```{r}
step(model_2_red, direction = "both")
```
Only site name and month are retained, using Anacostia NE and April as baselines

```{r}
model_fin <- lm(nitrogen_dioxide_no ~ SITE_NAME + Month, data = model_2_data)
summary(model_fin)
```


```{r}
plot(model_fin$resi)
```


```{r}
qqnorm(model_fin$resi)
qqline(model_fin$resi)
```


Residual plots indicate that linear model is appropriate.

```{r}
dwtest(model_fin)
```
Durbin Watson test reveals autocorrelation in this model



```{r}
n <- 119 # sample size
p <- 2 # 5 parameters

par(mfrow = c(2, 2))
plot(rstudent(model_fin), ylab = 'Studentized Deleted Residual')
abline(qt(1-0.1/(2*n), df = n - p - 1), 0, lty = 2)
abline(qt(0.1/(2*n), df = n - p - 1), 0, lty = 2)

plot(dffits(model_fin), ylab = 'DF Fits')
abline(1, 0, lty = 2)
abline(1, 0, lty = 2)

plot(cooks.distance(model_fin), ylab = "Cook's Distance")
abline(qf(0.5, df1 = 5, df2 = n - p), 0, lty = 2)

plot(hatvalues(model_fin), ylab = 'Leverege')
abline(2*p/n, 0, lty = 2)

vif(model_fin)

influenceIndexPlot(model_fin)
```

## Model 3: Does weather have an affect?

```{r}
model_3_data <- na.omit(subset(reduced_data, 
                               select = c("Season", "Month", "Year", 
                                          "SITE_NAME", "nitrogen_dioxide_no", "carbon_monoxide","outdoor_temperature", "relative_humidity"	, 
                                          "wind_direction__resultant", 
                                          "wind_speed__resultant")))
```

```{r}
model_3 <- lm(nitrogen_dioxide_no ~ ., data = model_3_data)
summary(model_3)
```

```{r}
step(model_3, direction = "both")
```

```{r}
model_3_step <- lm(formula = nitrogen_dioxide_no ~ Month + Year + SITE_NAME + 
    carbon_monoxide + outdoor_temperature + relative_humidity + 
    wind_direction__resultant + wind_speed__resultant, data = model_3_data)
summary(model_3_step)
```

## ANOVA test for significance of weather predictors
```{r}
model_3_red <- lm(formula = nitrogen_dioxide_no ~ Month + Year + SITE_NAME + 
    carbon_monoxide, data = model_3_data)

anova(model_3_red, model_3_step)
```
All weather variables cannot be dropped from the model, weather has some impact on concentration. 

## Ridge Regression to account for multicollinearity
```{r}
x_var <- model.matrix(model_3_step)
y_var <- as.numeric(model_3_data[[ "nitrogen_dioxide_no"]])
lambda_seq <- 10^seq(2, -1, by = -0.1)

ridge <- glmnet(x_var, y_var, alpha = 0, lambda = seq(0, 10, 0.01))
plot(ridge)

ridge_cv <- cv.glmnet(x_var, y_var, alpha = 0, lambda = seq(0, 10, 0.01))
best_lambda <- ridge_cv$lambda.min

ridge_out <- glmnet(x_var, y_var, alpha = 0, lambda = best_lambda)
summary(ridge_out)
coef(ridge_out)

predict(ridge, ridge_cv$lambda.min, newx = x_var, type = 'coefficients')

plot(ridge_cv)

y_predicted <- predict(ridge_out, s = best_lambda, newx = x_var)
sst <- sum((y_var - mean(y_var))^2)
sse <- sum((y_predicted - y_var)^2)
rsq <- 1 - sse/sst
rsq

best_lambda
```

